Pattern matching in\\ $(213,231)$-avoiding permutations

### 2015-10-08 article.tex

	A few particular cases of the pattern matching problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	(The algorithm also improves on the previous
	$O(|\pi| \log \log |\pi|)$ bound).
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear-time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns matching has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.
	The pattern matching problem is also solvable in
	polynomial-time for separable patterns \cite{Ibarra:1997,Bose:Buss:Lubiw:1998}
	(see also \cite{Bouvel:Rossin:Vialette:CPM:2007} for LCS-like issues
	of separable permutations).
	Separable permutations are those permutations that match neither
	$2413$ nor $3142$, and they are enumerated by the Schr{\"o}der numbers
%	(sequence A006318 in OEIS).
%	To the best of our knowledge,
%	separable permutations first arose in the work of
%	Avis and Newborn~\cite{Avis:Newborn:1981},
%	who showed that they are precisely the permutations which can be sorted by an
%	arbitrary number of pop-stacks in series,
%	where a pop-stack is a restricted form of stack in which any pop operation
%	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)

@@

	We focus in this paper on pattern matching issues for
	$(213,231)$-avoiding permutations
    (\emph{i.e.}, those permutations that avoid both $213$ and $231$).
	The number of $n$-permutations that avoid both
	$213$ and $231$ is
	$t_0 = 1$ for $n = 0$ and
	$t_n =2^{n-1}$ for $n\geq 1$ \cite{Simion:Schmidt:EJC:1985}.
	On an individual basis,
	the permutations that do not match the permutation pattern $231$
	are exactly the \emph{stack-sortable permutations} and they are counted by
	the Catalan numbers \cite{Knuth:1997:ACP:260999}.
	A stack-sortable permutation is a permutation whose elements may be sorted by
	an algorithm whose internal storage is limited to a single stack data structure.
	As for $213$, it is well-known that
 	if $\pi = \pi_1\pi\,\ldots\,\pi_n$ avoids $132$, then its complement
 	$\pi' = (n+1-\pi_1)(n+1-\pi_2)\,\ldots\,(n+1-\pi_n)$ avoids $312$, and
 	the reverse of $\pi'$ avoids $213$.
 	From a combinatorial point of view,
 	B\'ona \cite{Bona:ElJC:2012}
 	showed the rather surprising fact that the cumulative number of
 	matching of the classical patterns $231$ and $213$ are the same on the
 	set of permutations avoiding $132$,
 	beside the pattern based statistics $231$ and $213$
 	do not have the same distribution on this set.
 	\emph{Almost avoidance} has also been considered;
 	a permutation $\pi$ almost avoids a set $X$ of permutations
 	if there is a way to remove a single element of $\pi$ to get a permutation
 	that avoids all elements in $P$
 	and $L_n(P)$ denotes the set of permutations of length $n$ that almost avoid
 	$P$.
 	It is shown in \cite{Griffiths:Smith:Warren:PMA:2011} that
 	$L_n(213, 231) = L_n(132, 231) = L_n(132, 312) = L_n(213, 312)$.

	This paper is organized as follows.
	In Section~\ref{section:Definitions} the needed definitions are presented.
	Section~\ref{section:both are (213,231)-avoiding} is devoted to presenting
	an online linear-time algorithm in case both 
	permutations are $(213,231)$-avoiding,
	whereas Section~\ref{section:sigma only avoid 231 and 213} focuses on the case
	where only the pattern is $(213,231)$-avoiding.
	In Section~\ref{section:bivincular} we give a polynomial-time algorithm
	for $(213,231)$-avoiding bivincular patterns.
	In Section~\ref{section:LCS} we consider the problem of finding the longest
	$(213,231)$-avoiding pattern in permutations.

@@

\begin{proof}
  The forward direction is obvious.
  We prove the backward direction by induction on the size of the pattern
  $\sigma$.
  The base case is a pattern of size $2$.
  Suppose that $\sigma = 12$ and thus $\bijection(\sigma) = ascent$.
  Let $t = \pi_{i_1}\pi_{i_2}$, $i_1 < i_2$, be a subsequence of $\pi$
  such that $\bijection(t) = ascent$, this reduces to saying that
  $\pi_{i_1} < \pi_{i_2}$, 
  and hence that $t$ is a matching of $\sigma = 12$ in $\pi$.
  A similar argument shows that the lemma holds true for $\sigma = 21$.
  Now, assume that the lemma is true for all patterns up to size $k \geq 2$.
  Let $\sigma \in \Av{k+1}{231,213}$ and
  let $t$,
  be a subsequence of $\pi$ of length $k+1$ such that
  $\bijection(t) = \bijection(\sigma)$.
  As $\bijection(t)[2:] = \bijection(\sigma)[2:]$
  by the inductive hypothesis, it follows that
  $t[2:]$ is a matching of $\sigma[2:]$.
  Moreover $\bijection(t)[1] = \bijection(\sigma)[1]$ 
  thus $t[1]$ and $\sigma[1]$ are both either the minimal or the maximal
  element of their respective subsequences.
  Therefore, $t$ is a matching of $\sigma$ in $\pi$.
  \qed
\end{proof}

@@

\begin{proof}
According to Lemma~\ref{lemma:MatchStripeToPermutation} the problem reduces
to deciding whether $\bijection(\sigma)$ is a subsequence of $\bijection(\pi)$.
A natural greedy approach solves this issue in linear-time.
\qed
\end{proof}

@@

\begin{lemma}
Let $\ppattern$ be a permutation, 
$\factor(i)$ be an ascent (respectively descent) factor,
$s$ be a subsequence of $\pi$ such that $s \in \SET{\ppattern}{\ptext}{i}{j}$ and
that minimizes (resp. maximizes) the match of the left most element of $\factor(i-1)$.
For all subsequences 
$s' \in \SET{\ppattern}{\ptext}{i}{j}$ and for all subsequences $t$ of $\ptext$, such that $t=t's'$, 
if $t$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ such that the left most element of $\factor(i)$ is matched to $\ptext[j]$ then the subsequence $t's$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ such that the left most element of $\factor(i)$ is matched to $\ptext[j]$.
\end{lemma}

@@

\begin{proof}
By definition $s$ is a matching of $\ppattern[\LMEi(\factor(i)):]$. To prove that $ts$ is an matching of $\ppattern[\LMEi(\factor(i+1)):]$ we need to prove that every element of t is larger than every element of $s$. If $ts'$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ then every element of $t$ is larger than every element of $s'$. Moreover the maximal element of $s$ is smaller than  the maximal element of $s'$ so every element of $s$ is smaller than every element of $s'$ thus every element of $s$ is smaller than every element of $t$. We use a similar argument if $\factor(i)$ is a descent factor.
\qed
\end{proof}

@@

\begin{proof}
Suppose that there exists $l$, $i<l<j$,
such that $\ppattern[l]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[l]<\sigma[j]$ which is in contradiction with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
\qed
\end{proof}

@@

\noindent\textbf{BASE :} \\
$$
\PM{\sigma}{\ptext}{\lb}{\ub}{k}{j}=
\begin{cases}
	true 	& \text{if $\ptext[j] \in [\lb,\ub ]$}\\
			& \text{and if ${\ppattern[k]}_\lrcorner$ then $j=n$}\\
			& \text{and if ${\ppattern[k]}^\urcorner$ then $\ptext[j]=\ub=k$}\\
			& \text{and if  $^\ulcorner{\ppattern[k]}$ then $\ptext[j]=\lb=1$ } \\
			& \text{and if  $\overline{(\ppattern[k]-1)\ppattern[k] }$ then $\ptext[j]=\lb$ }  \\
			& \text{and if  $\overline{(\ppattern[k]\ppattern[k]-1)}$ then $\ptext[j]=\ub$}  \\

@@

\noindent\textbf{STEP :}

@@

	\item  If $\overline{\ppattern[i]\ppattern[i']}$,  (which implies that $\ppattern[i']=\ppattern[i]+1$) then if $\ppattern[i]$ is matched to $\ptext[j]$ then $\ppattern[i']$ must be matched to $\ptext[j]+1$.
		\begin{itemize}

@@

			\item If $\ppattern[i]$ is an ascent element and $\ppattern[i']$ is an ascent element then
			first remark that every recursive call
			$\PM{\sigma}{\ptext}{*}{\ub}{i'}{*}$ is true if $\sigma[i']$ is matched to element with value $\lb$. 
			Now remark that
			$i<i'$ and there is no  ascent element between $\ppattern[i]$ and $\ppattern[i']$ (lemma \ref{lemma:ascentDescentAscent}), 
			As a consequence $\sigma[i+1]$, $\sigma[i+2]$, $\ldots$, $\sigma[i'-1]$ are descent elements.  Moreover the recursive call from a descent element does not modified the lower bound and $\PM{\sigma}{\ptext}{\lb}{*}{i}{*}$ will put the lower bound to $\ptext[j]+1$ thus for every $\PM{\sigma}{\ptext}{\lb}{*}{i'}{*}$, $\lb=\ptext[j]+1$ (see figure $\ref{fig:ub dont change}$.

		\end{itemize}
\end{itemize}

@@

\section{Computing the longest $(213,231)$-avoiding subsequence}
\label{section:LCS}

@@

	\begin{proof}
	Let $s$ be a longest subsequence avoiding (213,231) with last element at index $f$ in $\pi$,
	suppose that $P(s)$ is not a longest increasing subsequence with last element at index $f$. Let $s_m$ be a longest increasing subsequence with last element $f$.
	Thus $|s_m|>|A(s)|$, clearly the sequence $s_m \cup D(s)$
	is $(213,231)$-avoiding and is longer than $s$, this is a contradiction.
	The same idea can be used to show that $D(\pi)$ is the longest decreasing subsequence.
	\qed
	\end{proof}

	\begin{proposition}
	Let $\pi$ be a permutation. One can compute
	the longest $(213,231)$-avoiding subsequence that can be matched in $\pi$
	in $O(n\log(\log(n)))$ time and in $O(n)$ space.
	\end{proposition}

	\begin{proof}
	The proposition \ref{proposition:longestIncreasingSubsequence} lead to an algorithm
	where one has to compute longest increasing and decreasing subsequence ending at every index. Then finding the maximal sum of longest increasing and decreasing subsequence ending at the same index.
	Computing the longest increasing subsequence and the longest decreasing subsequence can be done in 
	$O(n\log(\log(n)))$ time and $O(n)$ space 
	(see \cite{Bespamyatnikh00enumeratinglongest}), 
	then finding the maximal can be done in linear time.
	\qed
	\end{proof}

@@

\begin{proof}
Consider the following problem,
that compute the longest stripe common to $\pi_1$ and $\pi_2$.
Given two permutations $\pi_1$ and $\pi_2$,\\

$$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}
$$
\begin{center}
= $\max$ \{$|s|$ $|$ $s$ can be matched $\pi_1[i_1:]$ with every element of the match in $[\lb_1,\ub_1]$ and $s$ can be matched $\pi_2[i_2:]$ with every element of the match in $[\lb_2,\ub_2]$ \}
\end{center}

We show that we can compute by induction the problem.\\

\noindent\textbf{BASE :}

$$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{|\pi_1|}{|\pi_2|} =
\begin{cases}
	1 & \text{if $\lb_1 \leq \pi_1[j] \leq \ub_1$
	}\\
	& \text{ and $\lb_2 \leq \pi_2[j] \leq \ub_2$}\\
	0 & otherwise\\
\end{cases}
$$

\noindent\textbf{STEP :}

$$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}=max
\begin{cases}
	\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2+1} \\
	\\
	\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1+1}{i_2} \\
	\\
	\match{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}
\end{cases}
$$

@@

\bibliography{bibli}{}
\bibliographystyle{plain}
%

>>>>

	A few particular cases of the pattern matching problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	% (The algorithm also improves on the previous
	% $O(|\pi| \log \log |\pi|)$ bound.)
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear-time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns matching has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.
	The pattern matching problem is also solvable in
	polynomial-time for separable patterns \cite{Ibarra:1997,Bose:Buss:Lubiw:1998}
	(see also \cite{Bouvel:Rossin:Vialette:CPM:2007} for LCS-like issues
	of separable permutations).
	Separable permutations are those permutations that match neither
	$2413$ nor $3142$, and they are enumerated by the Schr{\"o}der numbers
%	(sequence A006318 in OEIS).
%	To the best of our knowledge,
%	separable permutations first arose in the work of
%	Avis and Newborn~\cite{Avis:Newborn:1981},
%	who showed that they are precisely the permutations which can be sorted by an
%	arbitrary number of pop-stacks in series,
%	where a pop-stack is a restricted form of stack in which any pop operation
%	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)

@@

	We focus in this paper on pattern matching issues for
	$(213,231)$-avoiding permutations
    (\emph{i.e.}, those permutations that avoid both $213$ and $231$).
	The number of $n$-permutations that avoid both
	$213$ and $231$ is
	$t_0 = 1$ for $n = 0$ and
	$t_n =2^{n-1}$ for $n\geq 1$ \cite{Simion:Schmidt:EJC:1985}.
	On an individual basis,
	the permutations that do not match the permutation pattern $231$
	are exactly the \emph{stack-sortable permutations} and they are counted by
	the Catalan numbers \cite{Knuth:1997:ACP:260999}.
	A stack-sortable permutation is a permutation whose elements may be sorted by
	an algorithm whose internal storage is limited to a single stack data structure.
	As for $213$, it is well-known that
 	if $\pi = \pi_1\pi\,\ldots\,\pi_n$ avoids $132$, then its complement
 	$\pi' = (n+1-\pi_1)(n+1-\pi_2)\,\ldots\,(n+1-\pi_n)$ avoids $312$, and
 	the reverse of $\pi'$ avoids $213$.
 	% From a combinatorial point of view,
 	% B\'ona \cite{Bona:ElJC:2012}
 	% showed the rather surprising fact that the cumulative number of
 	% occurrences of the classical patterns $231$ and $213$ are the same on the
 	% set of permutations avoiding $132$,
 	% beside the pattern based statistics $231$ and $213$
 	% do not have the same distribution on this set.
 	% \emph{Almost avoidance} has also been considered;
 	% a permutation $\pi$ almost avoids a set $X$ of permutations
 	% if there is a way to remove a single element of $\pi$ to get a permutation
 	% that avoids all elements in $P$
 	% and $L_n(P)$ denotes the set of permutations of length $n$ that almost avoid
 	% $P$.
 	% It is shown in \cite{Griffiths:Smith:Warren:PMA:2011} that
 	% $L_n(213, 231) = L_n(132, 231) = L_n(132, 312) = L_n(213, 312)$.
	This paper is organized as follows.
	In Section~\ref{section:Definitions} the needed definitions are presented.
	Section~\ref{section:both are (213,231)-avoiding} is devoted to presenting
	an online linear-time algorithm in case both 
	permutations are $(213,231)$-avoiding,
	whereas Section~\ref{section:sigma only avoid 231 and 213} focuses on the case
	where only the pattern is $(213,231)$-avoiding.
	In Section~\ref{section:bivincular} we give a polynomial-time algorithm
	for $(213,231)$-avoiding bivincular patterns.
	In Section~\ref{section:LCS} we consider the problem of finding the longest
	$(213,231)$-avoiding pattern in permutations.

@@

@@

@@

\begin{lemma}
\label{lemma:ts}
Let $\ppattern$ be a permutation, 
$\factor(i)$ be an ascent (respectively descent) factor,
$s$ be a subsequence of $\pi$ such that $s \in \SET{\ppattern}{\ptext}{i}{j}$ and
that minimizes (resp. maximizes) the match of the left most element of $\factor(i-1)$.
For all subsequences 
$s' \in \SET{\ppattern}{\ptext}{i}{j}$ and for all subsequences $t$ of $\ptext$, such that $t=t's'$, 
if $t$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ such that the left most element of $\factor(i)$ is matched to $\ptext[j]$ then the subsequence $t's$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ such that the left most element of $\factor(i)$ is matched to $\ptext[j]$.
\end{lemma}

@@

@@

@@

\noindent\textbf{BASE:} \\
$$
\PM{\sigma}{\ptext}{\lb}{\ub}{k}{j}=
\begin{cases}
	true 	& \text{if $\ptext[j] \in [\lb,\ub ]$}\\
			& \text{and if ${\ppattern[k]}_\lrcorner$ then $j=n$}\\
			& \text{and if ${\ppattern[k]}^\urcorner$ then $\ptext[j]=\ub=k$}\\
			& \text{and if  $^\ulcorner{\ppattern[k]}$ then $\ptext[j]=\lb=1$ } \\
			& \text{and if  $\overline{(\ppattern[k]-1)\ppattern[k] }$ then $\ptext[j]=\lb$ }  \\
			& \text{and if  $\overline{(\ppattern[k]\ppattern[k]-1)}$ then $\ptext[j]=\ub$}  \\

@@

\noindent\textbf{STEP:}

@@

	\item  If $\overline{\ppattern[i]\ppattern[i']}$,  (which implies that $\ppattern[i']=\ppattern[i]+1$) then if $\ppattern[i]$ is matched to $\ptext[j]$ then $\ppattern[i']$ must be matched to $\ptext[j]+1$.

		\begin{itemize}

@@

			\item If $\ppattern[i]$ is an ascent element and $\ppattern[i']$ is an ascent element then
			first remark that every recursive call
			$\PM{\sigma}{\ptext}{*}{\ub}{i'}{*}$ is true if $\sigma[i']$ is matched to element with value $\lb$. 
			Now remark that

			$i<i'$ and there is no  ascent element between $\ppattern[i]$ and $\ppattern[i']$ (lemma \ref{lemma:ascentDescentAscent}), 
			As a consequence $\sigma[i+1]$, $\sigma[i+2]$, $\ldots$, $\sigma[i'-1]$ are descent elements.  Moreover the recursive call from a descent element does not modified the lower bound and $\PM{\sigma}{\ptext}{\lb}{*}{i}{*}$ will put the lower bound to $\ptext[j]+1$ thus for every $\PM{\sigma}{\ptext}{\lb}{*}{i'}{*}$, $\lb=\ptext[j]+1$ (see figure $\ref{fig:ub dont change}$.
		\end{itemize}
\end{itemize}

@@

\section{Computing the longest $(213,231)$-avoiding pattern}
\label{section:LCS}

@@

	\begin{proposition}
	\label{proposition:longest 2}
	Let $\pi$ be a permutation. One can compute
	the longest $(213,231)$-avoiding subsequence that can be matched in $\pi$
	in $O(n\log(\log(n)))$ time and in $O(n)$ space.
	\end{proposition}

@@

\begin{proof}
Consider the following problem
that computes the longest stripe common to $\pi_1$ and $\pi_2$:
Given two permutations $\pi_1$ and $\pi_2$, we define
$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}
$
\begin{center}
= $\max$ \{$|s|$ $|$ $s$ can be matched $\pi_1[i_1:]$ with every element of the match in $[\lb_1,\ub_1]$ and $s$ can be matched $\pi_2[i_2:]$ with every element of the match in $[\lb_2,\ub_2]$ \}
\end{center}

We show how to solve this problem by dynamic programming.

\noindent\textbf{BASE:}
$$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{|\pi_1|}{|\pi_2|} =
\begin{cases}
	1 & \text{if $\lb_1 \leq \pi_1[j] \leq \ub_1$
	}\\
	& \text{ and $\lb_2 \leq \pi_2[j] \leq \ub_2$}\\
	0 & otherwise\\
\end{cases}
$$

\noindent\textbf{STEP:}
$$
\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}=max
\begin{cases}
	\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2+1} \\
	\\
	\LCS{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1+1}{i_2} \\
	\\
	\match{\pi_1}{\lb_1}{\ub_1}{\pi_2}{\lb_2}{\ub_2}{i_1}{i_2}
\end{cases}
$$

@@

\bibliography{bibli}{}
\bibliographystyle{plain}

\newpage
\section*{Appendix (Reviewers' version only)}

\begin{proof}[of Lemma~\ref{lemma:first element is 1 or n}]
Any other initial element would serve as a `$2$' in either a
$231$ or $213$ with $1$ and $n$ as the `$1$' and `$3$' respectively.
\qed
\end{proof}

\bigskip

\begin{proof}[of Lemma~\ref{lemma:MatchStripeToPermutation}]
  The forward direction is obvious.
  We prove the backward direction by induction on the size of the pattern
  $\sigma$.
  The base case is a pattern of size $2$.
  Suppose that $\sigma = 12$ and thus $\bijection(\sigma) = ascent$.
  Let $t = \pi_{i_1}\pi_{i_2}$, $i_1 < i_2$, be a subsequence of $\pi$
  such that $\bijection(t) = ascent$, this reduces to saying that
  $\pi_{i_1} < \pi_{i_2}$, 
  and hence that $t$ is a matching of $\sigma = 12$ in $\pi$.
  A similar argument shows that the lemma holds true for $\sigma = 21$.
  Now, assume that the lemma is true for all patterns up to size $k \geq 2$.
  Let $\sigma \in \Av{k+1}{231,213}$ and
  let $t$,
  be a subsequence of $\pi$ of length $k+1$ such that
  $\bijection(t) = \bijection(\sigma)$.
  As $\bijection(t)[2:] = \bijection(\sigma)[2:]$
  by the inductive hypothesis, it follows that
  $t[2:]$ is a matching of $\sigma[2:]$.
  Moreover $\bijection(t)[1] = \bijection(\sigma)[1]$ 
  thus $t[1]$ and $\sigma[1]$ are both either the minimal or the maximal
  element of their respective subsequences.
  Therefore, $t$ is a matching of $\sigma$ in $\pi$.
  \qed
\end{proof}

\bigskip

\begin{proof}
According to Lemma~\ref{Proposition:both permutations are avoiding} the problem reduces
to deciding whether $s_\sigma$ occurs as a subsequence in $s_\pi$.
A straightforward greedy approach solves this issue in linear-time.
\qed
\end{proof}

\bigskip

%\begin{proof}[of Lemma~\ref{lemma:whereIsMax}]
%Consider the rightmost element of $\factor{\sigma}{i}$ and assume this element is an ascent.
%Then, it follows that it is smaller than the following element which is $\first{\sigma}{i-1}$.
%Now, since $\factor{\sigma}{i}$ is an increasing subsequence, every element of
%$\factor{\sigma}{i}$ is smaller than $\first{\sigma}{i-1}$.
%According to Lemma~
%\ref{lemma:first element is 1 or n}, $\first{\sigma}{i-1}$ is 
%the maximal (minimal) element of $\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$, 
%and hence
%$\first{\sigma}{i-1}$ is the maximal  element of
%$\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$.
%
%A similar argument shows that the lemma holds true if the rightmost element
%of $\factor{\sigma}{i}$ is a descent.
%\qed
%\end{proof}

\bigskip

\begin{proof}[of Lemma~\ref{lemma:ts}]
By definition $s$ is a matching of $\ppattern[\LMEi(\factor(i)):]$. To prove that $ts$ is an matching of $\ppattern[\LMEi(\factor(i+1)):]$ we need to prove that every element of t is larger than every element of $s$. If $ts'$ is a matching of $\ppattern[\LMEi(\factor(i+1)):]$ then every element of $t$ is larger than every element of $s'$. Moreover the maximal element of $s$ is smaller than  the maximal element of $s'$ so every element of $s$ is smaller than every element of $s'$ thus every element of $s$ is smaller than every element of $t$. We use a similar argument if $\factor(i)$ is a descent factor.
\qed
\end{proof}

\bigskip

\begin{proof}[of Lemma~\ref{lemma:ascentDescentAscent}]
Suppose that there exists $l$, $i<l<j$,
such that $\ppattern[l]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[l]<\sigma[j]$ which is in contradiction with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
\qed
\end{proof}

\bigskip

\begin{proof}[of Proposition~\ref{proposition:longestIncreasingSubsequence}]
Let $s$ be a longest subsequence avoiding (213,231) with last element at index $f$ in $\pi$,
suppose that $P(s)$ is not a longest increasing subsequence with last element at index $f$. Let $s_m$ be a longest increasing subsequence with last element $f$.
Thus $|s_m|>|A(s)|$, clearly the sequence $s_m \cup D(s)$
is $(213,231)$-avoiding and is longer than $s$, this is a contradiction.
The same idea can be used to show that $D(\pi)$ is the longest decreasing subsequence.
\qed
\end{proof}

\bigskip

\begin{proof}[of Proposition~\ref{proposition:longest 2}]
The proposition \ref{proposition:longestIncreasingSubsequence} lead to an algorithm
where one has to compute longest increasing and decreasing subsequence ending at every index. Then finding the maximal sum of longest increasing and decreasing subsequence ending at the same index.
Computing the longest increasing subsequence and the longest decreasing subsequence can be done in 
$O(n\log(\log(n)))$ time and $O(n)$ space 
(see \cite{Bespamyatnikh00enumeratinglongest}), 
then finding the maximal can be done in linear time.
\qed
\end{proof}
